### 2014-06-24 tex/main-body-text.tex
new fig2 referenced in tex document
more minor edits

\subsection*{Analysis of coastal California steelhead} 
We used a published dataset of 2596 steelhead \textit{Oncorhynchus mykiss}
genotyped at 15 microsatellite loci from \citet{Garzaetal_norcal} to compare the programs
{\sc flock} and {\sc structure} .  {\sc flock} 
was run six times using the default setting for all 
run parameters (initial partition = random, number of iterations = 20, number of runs = 50, 
LLOD threshold = 0) for values of $K$ from 2 to 6. The stopping point of 
$K=6$ was chosen after four consecutive values of $K$
yielded no plateau (see below). For each value of $K$ from 2 to 6, the no-admixture, 
uncorrelated-allele-frequency model implemented in {\sc structure} was run 6 times with 
a 50,000 sweep burn-in followed by 150,000 sweep sample from the posterior. 
All runs were performed on a JNCS D685 with an i7 3.40GHz processor and 8GB
of RAM running 64-bit Windows 7. For each set of runs {\sc flock} reports the 
normalized likelihood values (the exponenetiated log-likelihood value for an individuals 
membership to each $K$ clusters scaled to sum to one) for the run that belongs to 
the longest plateau - the 'best run'. 
This value is comparable to the $q_i$ values
returned by {\sc structure}, which, in the no-admixture case, are posterior probabilities of cluster membership.
The program {\sc clumpp} \citep{Jak&Ros2007} was used to relabel clusters among different runs,
while the program {\sc district} \citep{Rosenberg2004} was used to plot individual \textit{$q_i$} values.

\citet{Duc&Tur2012} promote a method for using {\sc flock} to estimate, $K$, the number of
subpopulations. The user must first specify the number of runs (different initial allocation of individuals 
to each cluster) as well as the number of iterations each individual has the opportunity to be reallocated 
to a different cluster. For each run, the log-likelihood value of membership to all $K$ clusters is computed
and the log likelihood difference (LLOD), the difference between the cluster with the highest likelihood 
and second highest likelihood, is reported for each individual. Runs that converge on the same solution 
are indicated by the same mean LLOD score and 
are referred to as a plateau. The plateau record is a list of how many runs ended up at these different solutions. 
\citet{Duc&Tur2012} developed two stopping rules that lead to the estimation of $K$. In the abscence of 
reasonable values for $k$, \citet{Duc&Tur2012} advocate running  {\sc flock} until either a single plateau 
$\geq 6$  is reached or four consecutive $K$ values yield no plateaus. If the first stopping condition is satisfied,
the $K$ for which there is a single plateau $\geq 6$ is the point estimate for $K$. Alternatively, 
if multiple plateaus are observed with at least one plateau $\geq 6$, the largest value of $K$ that satisfies 
this conditon should be taken as the lower bound estimate of $K$. If no plateaus $\geq 6$ are observed, $K$
is \textit{undecided}.

@@

\subsection*{Results} 
Run times were relatively short for both programs. {\sc flock} 
required 1133 minutes to complete all the runs, averaging 188.8 minutes for 
each of six groups of runs over $K$ from 2 to 6. All runs of {\sc structure} were completed in 
486 minutes. Each of the six groups of runs averaged 80.8 minutes. Increasing
$K$ from of 2 to 6 increased the run time for {\sc flock} by a factor of 2.4, while the run time 
for {\sc structure} increased by a factor of only 1.6.
\hrule

The allocation of indivividuals into K clusters was similar for each program. For all six groups 
of runs from \textit{k} = 2-4, {\sc flock} resulted in nearly identical clustering of individuals 
(See supplementary materials) none of which differed substantially from results obtain from 
{\sc structure} (Fig 1). Of the 'best runs' for the six groups for \textit{k} of 2, two  
runs yielded identical clustering (same mean LLOD score), all other runs 
differed by at most the allocation of 13 individuals. As 
\textit{k} increases so do the number of solutions
that {\sc flock} finds (Fig 1). For all groups of runs for $K$ $\geq 2$, no plateaus were observed - all intial 
allocations resulted in solutions that differed. These solutions, however, are suprisingly similar to
 the partions identified by {\sc structure}. Both programs idenfity clusters 
which can be attributed to the geographical boundaries identified by \citet{Garzaetal_norcal}.

Despite the similar allocation of individuals into \textit{k} clusters, the 
plateau analysis to estimate $K$ in {\sc flock} leads to different conclusions among groups of runs, as well as 
one that differs from {\sc structure}. While no plateaus were observed in any of the six groups 
for a \textit{k} $\geq$2, the plateau record varied greatly among each
group for \textit{k}=2 (plateau  sequences- Run1= 3,2,3,7,4; Run2=2,5,2,3,2,3; Run3=2,2,2,2,2,5; 
Run4=2,3,4,2,4,3; Run5=2,2,5,5,3; and; Run6=2,2,2,2,2,3,3). The plateau analysis 
procedure suggested by \citet{Duc&Tur2012} leads to 
either an estimate of K$\geq$2 (Run1) or \textit{undecided} as a consequence of no plateau 
lengths being greater than 6. Interpreting the results from {\sc structure} is often accomplished by 
looking at the estimated log likelihood of the data given K, the delta k method \citep{Evannoetal2005}, and plots of 
individual \textit{$q_i$} values. Both the estimated likelihood of the data
and the plots of individuals \textit{$q_i$} values suggest that 
the most likely \textit{k} for the dataset is 6.

@@

\begin{figure*}
\begin{center}
	% NOTE. Pat made Q-val-StrFloc.pdf that has all runs individually, replaced with 
	% all of them on same axis. It will look fine as a black and white with different 
	% characters.
 \includegraphics[width=\textwidth]{images/Figures-Pat/Qval-1plotGreyScale.pdf}
    \caption{Individual \textit{$q_i$} values plotted for each cluster for \textit{k}=5. 
{\sc flock} estimates of likelihood of assignment to each cluster have less uncertainty
 attributed to them.}
    \label{Fig.2}
\end{center}
\end{figure*}

>>>>

\subsection*{Analysis of coastal California steelhead} 
We used a published dataset of 2596 steelhead \textit{Oncorhynchus mykiss}
genotyped at 15 microsatellite loci from \citet{Garzaetal_norcal} to compare the programs
{\sc flock} and {\sc structure} .  {\sc flock} 
was run six times using the default setting for all 
run parameters (initial partition = random, number of iterations = 20, number of runs = 50, 
LLOD threshold = 0) for values of $K$ from 2 to 6. The stopping point of 
$K=6$ was chosen after four consecutive values of $K$
yielded no plateau (see below). For each value of $K$ from 2 to 6, the no-admixture, 
uncorrelated-allele-frequency model implemented in {\sc structure} was run 6 times with 
a 50,000 sweep burn-in followed by 150,000 sweep sample from the posterior. 
All runs were performed on a JNCS D685 with an i7 3.40GHz processor and 8GB
of RAM running 64-bit Windows 7. For each set of runs {\sc flock} reports the 
normalized likelihood values (the exponenetiated log-likelihood value for an individuals 
membership to each $K$ clusters scaled to sum to one) for a run belonging to the longest plateau.
This value is comparable to the $q_i$ values
returned by {\sc structure}, which, in the no-admixture case, are posterior probabilities of cluster membership.
The program {\sc clumpp} \citep{Jak&Ros2007} was used to relabel clusters among different runs,
while the program {\sc district} \citep{Rosenberg2004} was used to plot individual \textit{$q_i$} values.

\citet{Duc&Tur2012} promote a method for using {\sc flock} to estimate, $K$, the number of
subpopulations. The user must specify both the number of runs (different initial allocation of individuals 
to each cluster) as well as the number of iterations each individual has the opportunity to be reallocated 
to a different cluster. For each run, the log-likelihood value for membership to all $K$ clusters is computed
and the log likelihood difference (LLOD), the difference between the cluster with the highest likelihood 
and second highest likelihood, is calculated for each individual. Runs that converge on the same solution 
are indicated by the same mean LLOD score and 
are referred to as a plateau. The plateau record is a list of how many runs converged at the same allocation
of individuals. 
\citet{Duc&Tur2012} developed two stopping rules that lead to the estimation of $K$. In the abscence of 
reasonable values for $k$, \citet{Duc&Tur2012} advocate running  {\sc flock} until either a single plateau 
$\geq 6$  is reached or four consecutive $K$ values yield no plateaus. If the first stopping condition is satisfied,
the $K$ for which there is a single plateau $\geq 6$ is the point estimate for $K$. Alternatively, 
if multiple plateaus are observed with at least one plateau $\geq 6$, the largest value of $K$ that satisfies 
this conditon should be taken as the lower bound estimate of $K$. If no plateaus $\geq 6$ are observed, $K$
is \textit{undecided}.

@@

\subsection*{Results} 
Run times were relatively short for both programs. All runs of {\sc flock} 
completed in 1133 minutes. Each of the six groups of runs averaged 188.8 minutes.
All runs of {\sc structure} were completed in 
486 minutes and averaged 80.8 minutes to complete each of the six groups of runs. Increasing
$K$ from of 2 to 6 increased the run time for {\sc flock} by a factor of 2.4, while the run time 
for {\sc structure} increased by a factor of only 1.6.
\hrule

The allocation of indivividuals into K clusters was similar for each program. For all six groups 
of runs from \textit{k} = 2 - 4, {\sc flock} resulted in nearly identical clustering of individuals 
(See supplementary materials) none of which differed substantially from results obtain from 
{\sc structure} (Fig 1). Of the 'best runs' for the six groups for \textit{k} of 2, two  
runs yielded identical clustering (same mean LLOD score), all other runs 
differed by at most the allocation of 13 individuals. As 
\textit{k} increases so do the number of solutions
that {\sc flock} finds (Fig 1). For all groups of runs for $K$ $\geq 2$, no plateaus were observed - all intial 
allocations resulted in solutions that differed. These solutions, however, are suprisingly similar to
 the partions identified by {\sc structure}. Both programs idenfity clusters 
which can be attributed to the geographical boundaries identified by \citet{Garzaetal_norcal}.

Despite the similar allocation of individuals into \textit{k} clusters, the 
plateau analysis to estimate $K$ in {\sc flock} leads to different conclusions among groups of runs, as well as 
one that differs from {\sc structure}. While no plateaus were observed in any of the six groups 
for a \textit{k} $\geq$2, the plateau record varied greatly among each
group for \textit{k} = 2 (plateau  sequences- Run1 = 3,2,3,7,4; Run2 = 2,5,2,3,2,3; Run3 = 2,2,2,2,2,5; 
Run4 = 2,3,4,2,4,3; Run5 = 2,2,5,5,3; and; Run6 = 2,2,2,2,2,3,3). The plateau analysis 
procedure suggested by \citet{Duc&Tur2012} leads to 
either an estimate of K $\geq$ 2 (Run1) or \textit{undecided} as a consequence of no plateau 
lengths being greater than 6. Interpreting the results from {\sc structure} is often accomplished by 
looking at the estimated log likelihood of the data given K, the delta k method \citep{Evannoetal2005}, and plots of 
individual \textit{$q_i$} values. Both the estimated likelihood of the data
and the plots of individuals \textit{$q_i$} values suggest that 
the most likely \textit{k} for the dataset is 6.

@@

\begin{figure*}
\begin{center}
	% NOTE. Pat replaced old Q-val-StrFloc.pdf with a ggplot version 
 \includegraphics[width=25em]{images/Figures-Pat/Fig2.pdf}
    \caption{Individual \textit{$q_i$} values plotted for each cluster for \textit{k}=5. 
{\sc flock} estimates of likelihood of assignment to each cluster have less uncertainty
 attributed to them.}
    \label{Fig.2}
\end{center}
\end{figure*}
